comgen-outline

### 2017-07-03 paper/comgen-paper.Rmd
reorganize habituals analysis section to provide parallel to generics; codify some numbers

From the prevalence prior data (Expt. 1a), we estimate participants' beliefs about the prevalence of a property for the target categories (e.g., the percentage of \textsc{robins} that \textsc{lay eggs}.
We find a little over half of the variance in the endorsement data is explained this way ($r^2(`r r2.gen.n`) = `r r2.gen.regression.prev`$; MSE=$`r mse.gen.regression.prev`$; Figure \ref{fig:commongenerics}b, right). 
This is because our stimulus set includes generics that are true with high-prevalence properties (e.g., *Leopards have spots.*) and generics that are false with low prevalence properties (e.g., *Leopards have wings.*). 
However, large deviations from an account based purely on target-category prevalence remain: Generics in which the target-category has intermediate prevalence (prevalence quartiles 2 and 3: $`r round(100*intermediate.prev.quantiles[["25%"]])`\% < \text{prevalence} < `r  round(100*intermediate.prev.quantiles[["76%"]])`\%$), are not at all explained by prevalence within those categories ($r_{Q2,3}^2(`r r2.gen.n.intermedprev`) = `r r2.gen.regression.prev.intermedprev`$; MSE = $`r mse.gen.regression.prev.intermedprev`$).

@@

Figure \ref{fig:priorScatter} (right) shows example inferred priors.

Some items show substantial differences between the genders (e.g., \emph{wears a bra}) and some show subtle differences (e.g., \emph{watches professional football}).
Recall that the prior distributions $P(h)$ used in the RSA are with respect to an alternative class of entities (e.g., other individuals). 
It's possible that when evaluating habitual statements under certain contexts, $P(h)$ is with respect to either *all people* or *people of the same gender*. 
For example, are the frequency conditions by which a man would qualify to "watch  football" (habitually) different than those by which a woman would qualify to "watch  football"?
We explore the possibility of different truth conditions for habituals of different gendered characters in Experiment 2b, for select items with priors that differ substantially by gender.

@@

```{r}
d.gen.endorse.regression.prevalence <- left_join(
  d.gen.endorse.bayes,
  d.gen.endorse.prev.cue.uniq %>%
    mutate(
      prediction_mean = predict(glm.gen.endorse.prev, ., type = "response"),
      prediction_ci_lower = predict(glm.gen.endorse.prev, d.gen.endorse.prev.cue.uniq.lower, type = "response"),
      prediction_ci_upper = predict(glm.gen.endorse.prev, d.gen.endorse.prev.cue.uniq.upper, type = "response")
    ) %>%
    mutate(sentence = paste(Category, " ", Property, ".", sep = ""))
) %>% mutate(
  sqErr = (MAP_h - prediction_mean) ^ 2
)

d.gen.endorse.regression <- bind_rows(
  d.gen.endorse.regression.prevalence.cuevalidity %>%
    mutate(src = "regression_Prev_Cuevalidity"),
  d.gen.endorse.regression.prevalence %>% 
    mutate(src = "regression_Prev")
  )
```

@@

@@

@@

#### Descriptive behavioral results

On each trial of the experiment, the participant was told a person did a particular past frequency and was asked to evaluate the corresponding habitual statements.
A simple null hypothesis is that habitual statements convey that the a person *has done this action before*, analagous to how the quantifier *Some* conveys that there exists at least one entity with the feature.
Under this hypothesis, all of the items used in our experiment would qualify as good habitual statements as the target character always did the action 3 times during some time window. 
A more sophisticated alternative hypothesis is that the degree to which the habitual statement is good is the frequency itself. 
<!-- Our computational model predicts that the the degree to which the habitual statement is good is the frequency of event, normalized by the alternative possible frequencies under the prior. -->
<!-- [MHT: SENTENCE ABOVE IS THE SOFT-SEMANTICS INTERPRETATION] -->

Both of the simple alternative hypotheses can immediately be ruled by out by visual inspection of Figure \ref{fig:habituals-endorsement-figure} (top-left), which shows the correspondence between the frequency of the event (transformed to a log times-per-year scale) and the human-judged felicity of the corresponding habitual sentence. 
First, it is clear that not all habitual statements are strongly assented to, even though it is true that all the stimuli involved an agent doing the action with some non-zero frequency (at least 3 times before).
It is also clear that a habitual sentence can receive strong agreement even when the actions are very infrequent (3 times in a 5-year interval; e.g. \emph{writes novels}, \emph{climbs mountains}), arguing against the second alternative.
We see in addition that even when actions are done relatively frequently (e.g. 3 times in a one month interval), some habitual sentences receive less than full endorsement (e.g. \emph{wears socks}, \emph{drinks coffee}). 
In our data, actions completed with a high frequency (3 times in a one week interval) receive at a minimum 75\% endorsement, though there is still variability among them (e.g. between 10-25\% of people disagree with \emph{wears a watch} and \emph{wears a bra}).
Overall, the raw log-frequency of action predicts only a fraction of the variability in responses ($r^2(93) = 0.33$).
For actions that are done on the time scale of years or longer (lower median of frequency), frequency itself no longer explains the endorsements ($r^2(50) = 0.07$)

@@

For our six items that showed substantial variability in the elicited priors between genders, we explore differences in endorsements. 
Differences in endorsements could suggest that the *comparison class* of entities in the prior distribution used in our model might be gender-specific (i.e., when evaluating "Mary drinks beer." participants implicitly consider the distribution of frequencies for other women and not other men).
Endorsement probabilities and 95\% credible intervals for these six items at each time interavl for each gender are showed in Figure \ref{fig:habituals-endorsement-gender-exploration-scatter}.
We see no appreciable differences in endorsements for the two genders. 
This suggests that, at least the minimal experimental contexts we are exploring, the comparison class is not different for female vs. male target characters.

#### Model criticism and comparison

>>>>

From the prevalence prior data (Expt. 1a), we estimate participants' beliefs about the prevalence of a property for the target categories (e.g., the percentage of \textsc{robins} that \textsc{lay eggs}.
We find a little over half of the variance in the endorsement data is explained this way ($r^2(`r r2.gen.n`) = `r r2.gen.regression.prev`$; MSE=$`r mse.gen.regression.prev`$; Figure \ref{fig:generics-endorsement-figure}, upper-right: lower-left facet). 
This is because our stimulus set includes generics that are true with high-prevalence properties (e.g., *Leopards have spots.*) and generics that are false with low prevalence properties (e.g., *Leopards have wings.*). 
However, large deviations from an account based purely on target-category prevalence remain: Generics in which the target-category has intermediate prevalence (prevalence quartiles 2 and 3: $`r round(100*intermediate.prev.quantiles[["25%"]])`\% < \text{prevalence} < `r  round(100*intermediate.prev.quantiles[["76%"]])`\%$), are not at all explained by prevalence within those categories ($r_{Q2,3}^2(`r r2.gen.n.intermedprev`) = `r r2.gen.regression.prev.intermedprev`$; MSE = $`r mse.gen.regression.prev.intermedprev`$).

@@

Figure \ref{fig:priorScatter} (right) shows example inferred priors.
Some items show substantial differences between the genders (e.g., \emph{wears a bra}) and some show subtle differences (e.g., \emph{watches professional football}).
Recall that the prior distributions $P(h)$ used in the RSA are with respect to an alternative class of entities (e.g., other individuals). 
It's possible that when evaluating habitual statements under certain contexts, $P(h)$ is with respect to either *all people* or *people of the same gender*. 
For example, are the frequency conditions by which a man would qualify to "watch  football" (habitually) different than those by which a woman would qualify to "watch  football"?
We explore the possibility of different truth conditions for habituals of different gendered characters in Experiment 2b, for select items with priors that differ substantially by gender.

@@

@@

r2.hab.n <- length(r2.habituls.regression.freq$MAP_h)

r2.habituls.regression.freq <- compute_r2(d.hab.endorse.regression.freq,
                                          "MAP_h", "prediction")

mse.habituls.regression.freq <- compute_mse(d.hab.endorse.regression.freq,
                                          "MAP_h", "prediction")

r2.habituls.regression.infrequent.freq <- d.hab.endorse.regression.freq %>%
  filter(logAnnualRate < 1.1)

r2.hab.infrequent.n <- length(r2.habituls.regression.infrequent.freq$MAP_h)

r2.habituls.infrequent.regression.freq <- compute_r2(r2.habituls.regression.infrequent.freq,
                                          "MAP_h", "prediction")

mse.habituls.infrequent.regression.freq <- compute_mse(r2.habituls.regression.infrequent.freq,
                                          "MAP_h", "prediction")

@@

r2.habituls.regression.freq.distinct <- compute_r2(d.hab.endorse.regression.freq.mix,
                                          "MAP_h", "prediction")

mse.habituls.regression.freq.distinct <- compute_mse(d.hab.endorse.regression.freq.mix,
                                          "MAP_h", "prediction")

@@

Parallel to our analysis of generic language endorsements, we articulate a set of simple regression models and an alternative to our uncertain-threshold RSA model in order to understand the data and our model.
Before comparing models, we perform an exploratory analysis on a subset of our data, exploring whether or not endorsements for items whose priors differed by gender will also differ.

For our six items that showed substantial variability in the elicited priors between genders, we explore differences in endorsements. 
Differences in endorsements could suggest that the *comparison class* of entities in the prior distribution used in our model might be gender-specific (i.e., when evaluating "Mary drinks beer." participants implicitly consider the distribution of frequencies for other women and not other men).
Endorsement probabilities and 95\% credible intervals for these six items at each time interavl for each gender are showed in Figure \ref{fig:habituals-endorsement-gender-exploration-scatter}.
We see no appreciable differences in endorsements for the two genders. 
This suggests that, at least the minimal experimental contexts we are exploring, the comparison class is not different for female vs. male target characters.
For all remaining analyses, we collapse across character gender when analyzing endorsements.

#### Frequency baseline

The simplest alternative hypothesis for habitual language endorsement is that the degree to which the habitual statement is good is the frequency itself. 
This is unlikely to be the case based on Figure \ref{fig:habituals-endorsement-figure} (top-left), which shows the correspondence between the frequency of the event (transformed to a log times-per-year scale) and the human-judged felicity of the corresponding habitual sentence. 
First, it is clear that not all habitual statements are strongly assented to, even though it is true that all the stimuli involved an agent doing the action with some non-zero frequency (at least 3 times before).
It is also clear that a habitual sentence can receive strong agreement even when the actions are very infrequent (3 times in a 5-year interval; e.g. \emph{writes novels}, \emph{climbs mountains}), arguing against the second alternative.
We see in addition that even when actions are done relatively frequently (e.g. 3 times in a one month interval), some habitual sentences receive less than full endorsement (e.g. \emph{wears socks}, \emph{drinks coffee}). 
In our data, actions completed with a high frequency (3 times in a one week interval) receive at a minimum 75\% endorsement, though there is still variability among them (e.g. between 10-25\% of people disagree with \emph{wears a watch} and \emph{wears a bra}).
Overall, the raw log-frequency of action predicts only a fraction of the variability in responses ($r^2(`r r2.hab.n`) = `r r2.habituls.regression.freq`$; MSE=$`r mse.habituls.regression.freq`$). 
In addition, for actions that are done on the time scale of years or longer (lower median of frequency), frequency itself no longer explains the endorsements ($r^2(`r r2.hab.infrequent.n`) = `r r2.habituls.infrequent.regression.freq`$; MSE =$`r mse.habituls.infrequent.regression.freq`$)

@@

#### Frequency and distinctiveness

FILL IN

#### Fixed-threshold RSA model

On each trial of the experiment, the participant was told a person did a particular past frequency and was asked to evaluate the corresponding habitual statements.
A simple null hypothesis is that habitual statements convey that the a person *has done this action before*, analagous to how the quantifier *Some* conveys that there exists at least one entity with the feature.
Under this hypothesis, all of the items used in our experiment would qualify as good habitual statements as the target character always did the action 3 times during some time window.

#### Uncertain-threshold RSA model
